---
title: "Bike Sharing"
author: "Stefan Buchholz"
date: "Sunday, October 12, 2014"
output: html_document
---

Bike Sharing Stuff

## Read in the data
Firstly we load all the training and test data into its variables.

```{r read in the data}
# declare column types
train.columntypes <- c('factor', #have to correct that later
            'factor',
            'factor',
            'factor',
            'factor',
            'numeric',
            'numeric',
            'integer',
            'numeric',
            'integer',
            'integer',
            'integer')

test.columntypes <- c('factor', #this time its 3 columns less than in training dataset
                  'factor',
                  'factor',
                  'factor',
                  'factor',
                  'numeric',
                  'numeric',
                  'integer',
                  'numeric')

train.raw <- read.csv("train.csv",colClasses = train.columntypes)
test.raw <- read.csv("test.csv",colClasses = test.columntypes)
```

## Combine the data
Due to the fact that we want to have the same columns to work with we are about to add the missing columns to the test-dataset to join
it later on with the train dataset.

```{r combining the data}
test.raw$casual <- NA
test.raw$registered <- NA
test.raw$count <- NA

overalldata <- rbind(train.raw, test.raw) # split has to be made on row 10887: end
```
## Feature Engineering
Since we now have our overall data we can begin to add new features to the data.

```{r feature engineering}
#Convert the datetime into real datetime (german locale)
library(lubridate)
overalldata$datetime <- ymd_hms(overalldata$datetime)
overalldata$ger_datetime <- format(ymd_hms(overalldata$datetime), "%d.%m.%y %H:%M:%S")
#Feature Engineering

#extract the hour and weekday of the date
overalldata$hour <- hour(overalldata$datetime)
overalldata$year <- year(overalldata$datetime)
overalldata$month <- month(overalldata$datetime)
overalldata$weekday <- weekdays(overalldata$datetime)
overalldata$weekday <- as.factor(overalldata$weekday)
overalldata$combined <- paste(as.character(overalldata$weekday), "_", as.character(overalldata$hour), sep="")
overalldata$combined <- as.factor(overalldata$combined)
```

## Split back into train, test and an additional validation set
ok now we got our additional engineered features we have to split the overall data again in train, validation and test sets.

```{r splitting stuff}
full_train <- overalldata[1:10886,]

#constructing validation set
library(caTools)
set.seed(47533)
split = sample.split(full_train$count, SplitRatio = 0.75)
train = subset(full_train, split == TRUE)
validate = subset(full_train, split == FALSE)

test <- overalldata[10887:nrow(overalldata),]
```
## Modeling time
Now we can model and pick the best model and validate it on our validation set. We first select a model and second build two separate models for casual and registred because the count variable is constructed of it.

```{r modeling}
library(caret)
library(randomForest)

lm.model.casual.1 <- train(casual ~ season + holiday + workingday + weather + temp + atemp + humidity + windspeed + hour + as.factor(weekday), data=train, method="lm")

rpart.model.registered.1 <- train(registered ~ season + holiday + workingday + weather + temp + atemp + humidity + windspeed + year + month + weekday + hour, data=train, method="rpart")

rf.model.casual.1 <- randomForest(casual ~ season + holiday + workingday + weather + temp + atemp + humidity + windspeed + year + month + weekday + hour, data=train, method="rpart")

rf.model.registered.1 <- randomForest(registered ~ season + holiday + workingday + weather + temp + atemp + humidity + windspeed + year + month + weekday + hour, data=train, method="rpart")

gbm.casual.model <- train(casual ~ year + month + hour + weekday + season + holiday + workingday + weather + temp + atemp + humidity, data=train, method="gbm")

gbm.registered.model <- train(casual ~ year + month + hour + weekday + season + holiday + workingday + weather + temp + atemp + humidity, data=train, method="gbm")

#evalute it
library(qpcR)
RMSE(gbm.casual.model)

library(Metrics)
test.predictions <- predict(rf.model.registered.1, newdata=validate)
rmsle(validate$registered,test.predictions)
```
### Conclusion
#gbm (RMSE = 20.1 with shrinkage = 0,1 n.trees = 150 and interaction.depth = 3)
1. gbm = 19.73046 (Casual) / 2.935967
2. lm = 35.33 (casual) / 2.334489
3. rf = 38.49 (registered) / 0.4335096 <- best rmsle
4. rpart = 105.59 (registered) / 0.9594557

The winner is gbm!!!

## Predict the future
We got the best model (gbm) which we picked so lets predict the sharings for the future.

```{r predict the future}
#build final models (I had to use the log so that i dont get negative values when predicting)
gbm.casual.model <- train(log(casual +1) ~ year + month + hour + weekday + season + holiday + workingday + weather + temp + atemp + humidity, data=full_train, method="gbm")

gbm.registered.model <- train(log(registered +1) ~ year + month + hour + weekday + season + holiday + workingday + weather + temp + atemp + humidity, data=full_train, method="gbm")


#predict test casual and registered and combine it
casual.predictions <- predict(gbm.casual.model, newdata=test)
registered.predictions <- predict(gbm.registered.model, newdata=test)
predictions <- casual.predictions + registered.predictions
predictions <- exp(predictions)
summary(predictions)
hist(train$count)
hist(predictions)
################################

# Since gbm predicts negative values and i dont know how to fix this I try it with lm 
lm.model.casual.1 <- train(log(casual) ~ season + holiday + workingday + weather + temp + atemp + humidity + windspeed + hour + as.factor(weekday), data=train, method="lm")

lm.model.registered.1 <- train(log(registered) ~ season + holiday + workingday + weather + temp + atemp + humidity + windspeed + hour + as.factor(weekday), data=train, method="lm")

casual.predictions <- predict(lm.model.casual.1, newdata=test)
registered.predictions <- predict(lm.model.registered.1, newdata=test)
predictions <- casual.predictions + registered.predictions
predictions <- exp(predictions)

summary(predictions)

####################################
#Random forrest with best rmsle
rf.model.casual.1 <- randomForest(casual ~ season + holiday + workingday + weather + temp + atemp + humidity + windspeed + year + month + weekday + hour, data=full_train, method="rpart")

rf.model.registered.1 <- randomForest(registered ~ season + holiday + workingday + weather + temp + atemp + humidity + windspeed + year + month + weekday + hour, data=full_train, method="rpart")

casual.predictions <- predict(rf.model.casual.1, newdata=test)
registered.predictions <- predict(rf.model.registered.1, newdata=test)
predictions <- casual.predictions + registered.predictions

#1st try
predictions.rounded <- round(predictions)

#2nd try
predictions.ceiling <- ceiling(predictions)

summary(predictions.rounded)
summary(predictions.ceiling)
summary(validate$count)
```
## Construct submission file
construct a file for submission

```{r construct the submission file}
submit <- data.frame(datetime = test$datetime, count = predictions.ceiling)
write.csv(submit, file = "submission_rf_ceiling.csv", row.names = FALSE)
```